---
title: "Covid91 vaccine study Final2023"
author: "Dominic Garibay!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$; There is no effectiveness in the new Covid921 vaccine and the placebo vaccine when dealing with males

$H_A$; There is effectiveness in the new Covid921 vaccine and the placebo vaccine when dealing with males

```{r}
library(dplyr)
male = filter(FinalData,FinalData$sex=='M')
```

# Methods

Since we are comparing sex (male) to treatment, which is categorical~categorical, so we in turn will be using the Cat~Cat helper file

# Results

<!-- Split into Descriptive results and Inferential results. !--> 

## Descrpitive Results

<!-- Split into Graphical results and Numerical Results. !-->

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment ,data=male)
barchartGC(~infected + treatment ,data=male, type="percent")
```

From the graph we can see that, those who did not have covid did not experience any difference. Those who were infected with covid 19 and saw much better results than those with a placebo

### Numerical Descriptive Results

```{r}
table1 <- xtabs( ~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```

From the data we can see that the individuals with the vaccine have half of the infection rate, than those who had gotten the placebo.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
```
With the chisq test, we receive a p-value of 1.098e^-13, which is significantly smaller than 0.05. Through this test we can reject the null hypothesis of there being no difference in vaccine effectiveness in males.

```{r}
male$infected = factor(male$infected,c("OK","covid19"))
fisher.test(male$treatment,male$infected)
```

Through the fisher exact we receive a p-value of 4.631e^-14, which is also significantly less than 0.05. We also have .44 odds of getting infected with the vaccine.

# Discussion and Conclusion 

From our chisq test we receive a p-value of 1.098e^-13 and on our fisher exact test we receive a p-value of 4.631e^-14. From both of these test we can conclude that the vaccine is effective in males. The odds are 2 : 1  in favor of the vaccine for males.



## Females

$H_0$; There is no effectiveness in the new Covid921 vaccine and the placebo vaccine when dealing with females

$H_0$; There is effectiveness in the new Covid921 vaccine and the placebo vaccine when dealing with females

```{r}
library(dplyr)
female = filter(FinalData,FinalData$sex=='F')
```



## LGBTQ





## Druggies





# Overall Results and Conclusions



